\emph{Light Exposure Behavior Assessment (LEBA)}: Development of a novel instrument to capture light exposure-related behaviours

### 2021-11-12 manuscript.Rmd
Table 1 is formatted

```{r pacman, include=FALSE}
#You need to run this chunck 1X in your system 
#install.packages("pacman")
pacman::p_load(MOTE, tidyverse, psych, lavaan,kableExtra,gt,gtsummary, mirt,likert,kutils,semPlot,semTable,semTools,ggcorrplot,dlookr,qgraph,paran,EFA.MRFA,VIM,DiagrammeR,DiagrammeRsvg,ggplot2,cowplot,rsvg,questionr,magick, simsem)
pacman::p_install_gh("jthomasmock/gtExtras")
pacman::p_install_gh("crsh/papaja","crsh/citr", "masiraji/tabledown")
pacman::p_install_gh("masiraji/tabledown")

@@

library(papaja) # devtools::install_github("crsh/papaja")
library(lavaan)
library(semPlot) #devtools::install_github('SachaEpskamp/semPlot',  dependencies = T)
library(semTools)
library(MOTE)
library(car)#revercoding 
library(psych)
library(dlookr)
library(plyr)
library(tidyverse)
library(qgraph)
library(kableExtra)
library(paran) # Parallel analysis
library(EFA.MRFA) # Hull method
library(DiagrammeR) #devtools::install_github('rich-iannone/DiagrammeR')
library(DiagrammeRsvg) #For DiagrammeR
library(rsvg) #For DiagrammeR
library(ggcorrplot)
library(semTable)
library(magick)
library(mirt)
library(gtsummary)
library(gt)
library(gtExtras)
library(likert)
library(VIM) #Missing data
library(kutils)
library(simsem)
library(tabledown)

r_refs("references.bib")
```

@@

+-------------------------------------------------------------------+--------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------+-------------+
| Name                                                              | Authors                  | Description                                                                                                                                                  | Relevant Items                       | Adaptations |
+===================================================================+==========================+==============================================================================================================================================================+======================================+=============+
| Visual Light Sensitivity Questionnaire-8                          | [@verriotto2017new]      | Eight-question survey to assess the presence and severity of photosensitivity symptoms                                                                       | All items                            |             |
+-------------------------------------------------------------------+--------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------+-------------+
| Office Light Survey                                               | [@eklund1996development] | Multi-item questionnaire to assess electrical lighting environment in office                                                                                 |                                      |             |
+-------------------------------------------------------------------+--------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------+-------------+
| Harvard Light Exposure Assessment Questionnaire                   | [@bajaj2011validation]   | Self-administered semi-quantitative light questionnaire                                                                                                      | All Items                            |             |
+-------------------------------------------------------------------+--------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------+-------------+
| Hospital Lighting Survey                                          | [@dianat2013objective]   | 23 items questionnaire to assess light environment in a hospital                                                                                             |                                      |             |
+-------------------------------------------------------------------+--------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------+-------------+
| Morningness-Eveningness Questionnaire                             | [@horne1976self]         | 19 items questionnaire to understand your body clock                                                                                                         |                                      |             |
+-------------------------------------------------------------------+--------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------+-------------+
| Munich Chronotype Questionnaire (MCTQ)                            | [@roenneberg2003life]    | 17 items questionnaire to understand individuals phase of entrainment                                                                                        |                                      |             |
+-------------------------------------------------------------------+--------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------+-------------+
| Assessment of Sleep Environment                                   | [@Olivier.2016]       | 13 items questionnaire measuring your sleep environment quality                                                                                              |                                      |             |
+-------------------------------------------------------------------+--------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------+-------------+
| The Pittsburgh Sleep Quality Index (PSQI)                         | [@buysse1989pittsburgh]  | 9 items inventory to measure sleep quality and sleeping pattern                                                                                              |                                      |             |
+-------------------------------------------------------------------+--------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------+-------------+
| Self-Rating of Biological Rhythm Disorder for Adolescents (SBRDA) | [@Xie.2021]           | 29 items questionnaire assessing four dimensions of biological rhythm disorder in adolescents (digital media use, sleep, eating habits, and activity)        | Items 3., 22., 23., 24., 25. and 29. |             |
+-------------------------------------------------------------------+--------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------+-------------+
| Photosensitivity Assessment Questionnaire (PAQ)                   | [@Wu.2017]            | 16 dichotomous (yes/no) items questionnaire to assess "photophobia" and "photophilia", giving two final scores of "photophobic" and "photophilic" behaviours | All items                            |             |
+-------------------------------------------------------------------+--------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------+-------------+

: Existing related Scales

@@

-   Sleep disturbance and sleep-related impairment (adult and pediatric versions) [@Bevans.2019; @Buysse.2010; @Forrest.2018; @Harb.2015; @Yu.2011]
-   Sleep duration, timing, and latency, chronotype, social jetlag, time in bed, work/sleep schedule and outdoor light exposure duration (version for adults and adolescents) [@roenneberg2003life]
-   Sleep environment [@Olivier.2016\
-   Meal timing & caffeine consumption [custom items]
-   Light sensitivity (photophobia vs. photophilia) [@Wu.2017]
-   Self-reported pubertal stage (only if younger than 18 years old) [@Petersen.1988]

@@

Age among all participants ranged from `r min(desctable.data$slypos_demographics_age)` years to `r max(desctable.data$slypos_demographics_age)` years [EFA: *min* = `r min(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "1. EFA Sample"])`, *max* = `r max(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "1. EFA Sample"])`; CFA: *min* = `r min(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "2. CFA Sample"])`, *max* = `r max(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "2. CFA Sample"])`], with an overall mean of ~ `r round(mean(desctable.data$slypos_demographics_age), 0)` years of age [Overall: *M* = `r mean(desctable.data$slypos_demographics_age)`, *SD* = `r sd(desctable.data$slypos_demographics_age)`; EFA: *M* = `r mean(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "1. EFA Sample"])`, *SD* = `r sd(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "1. EFA Sample"])`; CFA: *M* = `r mean(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "2. CFA Sample"])`, *SD* = `r sd(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "2. CFA Sample"])`]. In total `r desc_tibble[3,2]` of the participants indicated female sex [EFA: `r desc_tibble[3,3]`; CFA: `r desc_tibble[3,4]`], `r desc_tibble[4,2]` indicated male [EFA: `r desc_tibble[4,3]`; CFA: `r desc_tibble[4,4]`] and `r desc_tibble[5,2]` indicated other sex [EFA: `r desc_tibble[5,3]`, CFA: `r desc_tibble[5,4]`]. Overall, `r desc_tibble[6,2]` [EFA: `r desc_tibble[6,3]`; CFA: `r desc_tibble[6,4]`] participants indicated a gender-variant identity. In a "Yes/No" question regarding native language, `r desc_tibble[7,2]` of respondents [EFA: `r desc_tibble[7,3]`; CFA: `r desc_tibble[7,4]`] indicated to be native English speakers. For their "Occupational Status", more than half of the overall sample reported that they currently work [Overall: `r desc_tibble[9,2]`; EFA: `r desc_tibble[9,3]`; CFA: `r desc_tibble[9,4]`], whereas `r desc_tibble[10,2]` [EFA: `r desc_tibble[10,3]`; CFA: `r desc_tibble[10,4]`] reported that they go to school and `r desc_tibble[11,2]` [EFA: `r desc_tibble[11,3]`; CFA: `r desc_tibble[11,4]`] responded that they do "Neither". With respect to the COVID-19 pandemic we asked participants to indicate their occupational setting during the last four weeks: In the overall sample `r desc_tibble[13,2]` [EFA: `r desc_tibble[13,3]`; CFA: `r desc_tibble[13,4]`] of the participants indicated that they were in a home office/ home schooling setting., while `r desc_tibble[14,2]` overall [EFA: `r desc_tibble[14,3]` ; CFA: `r desc_tibble[14,4]`] reported face-to-face work/schooling. Lastly, `r desc_tibble[15,2]` overall [EFA: `r desc_tibble[15,3]` ; CFA: `r desc_tibble[15,4]`] reported a combination of home- and face-to-face work/schooling, whereas `r desc_tibble[16,2]` overall [EFA: `r desc_tibble[16,3]`; CFA: `r desc_tibble[16,4]`] filled in the "Neither (no work or school, or indication)" response option. We tested all demographic variables in Table 1 for significant group differences between the EFA and CFA sample, applying Wilcoxon rank sum test for the continuous variable "Age" and Pearson's Chi-squared test for all other categorical variables via the gtsummary R package's "add_p" function [@sjoberg2021] . The p-values were corrected for multiple testing applying false discovery rate (FDR) via the "add_q" function of the same package. After p-value (FDR) correction for multiple testing, none of the demographic variables were significantly different between the EFA sample and the CFA sample (all q-values *q* ≥ `r min(desc_tibble$"**q-value**", na.rm=T)` , indicating equivalence).

@@

We used R (version 4.1.0), including several R packages, for our analyses.Initially, our tool have six poin Likert type response scale(0:Does not apply/I don't know; 1:Never, 2:Rarely; 3:Sometimes; 4:Often; 5: Alsways). As our purpose was to capture light exposure related behavior, "Does not apply/I don't know" and "Never" were providing similar information. As such we decided to collapse "Does not apply/I don't know" and "Never" options into one making it a 5 point Likert type response scale. Necessary assumptions of EFA, including sample adequacy, normality assumptions, quality of correlation matrix, were assessed. Our data violated both the univariate and multivariate normality assumptions. Due to these violations and the ordinal nature of our response data, we used a polychoric correlation matrix [@desjardinsHandbookEducationalMeasurement2018] for the EFA. We employed principal axis (PA) as a factor extraction method with varimax rotation. PA is robust to the normality assumption violations [@watkinsStepbyStepGuideExploratory2020]. The obtained latent structure was confirmed by another factor extraction method: the minimum residuals extraction method as well. We used a combination factor identification method including scree plot[@cattellScreeTestNumber1966], Horn's parallel analysis [@hornRationaleTestNumber1965], minimum average partials method[@velicerDeterminingNumberComponents1976], and hull method [@lorenzo-sevaHullMethodSelecting2011] to identify factor numbers. Additionally, to determine the simple structure, we followed the following guidelines recommended by psychometricians (i) no factors with fewer than three items (ii) no factors with a factor loading \<0.3 (iii) no items with cross-loading greater than .3 across factors [@bandalosFactorAnalysisExploratory2018] We also conducted psychometric analysis on non-merged response options data (supplementary analysis) and rejected the latent structure obtained as the factors were less interpretable.

@@

Sampling adequacy was checked using Kaiser-Meyer-Olkin (KMO) measures of sampling adequacy[@kaiserIndexFactorialSimplicity1974] . The overall KMO vale for 48 items was `r printnum(KMO$MSA)` which was above the cutoff value (.50) indicating a mediocre sample [@hutchesonMultivariateSocialScientist1999]. Table\@ref(tab:tabDes) summarizes the univariate descriptive statistics for the 48 items. some of the items were skewed with high Kurtosis values. Our data violated both univariate normality (Shapiro-Wilk statistics; [@shapiroAnalysisVarianceTest1965]) and multivariate normality assumptions (Marida's test;[@mardiaMeasuresMultivariateSkewness1970]). Multivariate skew was = `r printnum(mardia$b1p)` (p \<0.001) and multivariate kurtosis was = `r printnum(mardia$b2p)` (p \<0.001). Due to these violations and ordinal nature of the response data polychoric correlations over Pearson's correlations was chosen [@desjardinsHandbookEducationalMeasurement2018]. Bartlett's test of sphericity [@bartlettNoteMultiplyingFactors1954], $\chi^2$ (`r bartlet$df`) = `r apa(bartlet$chisq, 2, T)`, p \< .001\] indicated the correlations between items are adequate for the EFA. However only `r apa(cor.per, 2,T)`% of the inter-item correlation coefficients were greater than .30. The inter item correlation ranged between `r apa(min.cor,2,F)` to `r apa(max.cor,2,F)`. And the corrected item-total correlations ranged between `r apa(low.r.corec,2,F)` to `r apa(high.r.corec,2,F)`.

@@

With initial 48 items we conducted three rounds of EFA gradually discarded problematic items. (cross-loading items and poor factor loading (\<.30) items). Finally, a five-factor EFA solution with 25 items was accepted with low RMSR = 0.08 [@brownConfirmatoryFactorAnalysis2015], all factor-loading higher than .30 and no cross-loading greater than .30. We confirmed this five-factor latent structure using varimax rotation with a minimum residual extraction method (Table\@ref(tab:MinResTab)). Table\@ref(tab:EFATable) displays the factor-loading (structural coefficients) and communality of the items. The absolute value of the factor-loading ranged from `r apa(min.loadings,2,F)` to `r apa(max.loadings, 2, F)` indicating strong coefficients. The commonalities ranged between `r apa(min.com,2,F)` to `r apa(max.com, 2, F)`. However, the histogram of the absolute values of non-redundant residual-correlations Fig\@ref(fig:EFAResiduals) showed `r apa(sumR,0, T)`% correlations greater than the absolute value of .05, indicating a possible under-factoring. [@desjardinsHandbookEducationalMeasurement2018a]. Subsequently, we fitted a six-factor solution. However, a factor emerged with only one salient variable loading in the six-factor solution, thus disqualifying the six-factor solution (Table\@ref(tab:sixFacTab)).

In the five-factor solution, the first factor contained three items and explained `r apa(var1,2,F)`% of the total variance with a satisfactory internal reliability coefficient ($\alpha$ = `r apa(F1.alpha,2,F)`). All the items in this factor stemmed from the individual's preference to use blue light filters in different light environments. The second factor contained six items and explained `r apa(var2,2,F)`% of the total variance with a satisfactory internal reliability coefficient ($\alpha$ = `r apa(F2.alpha,2,F)`). Items under this factor commonly investigate an individual's hours spent outdoor. The third factor contained five items and explained `r apa(var3,2,F)`% of the total variance. Items under this factor dealt with the specific behaviors pertaining to sleep. The internal consistency reliability coefficient was, $\alpha$ = `r apa(F3.alpha,2,F)`. The fourth factor contained five items and explained `r apa(var4,2,F)`% of the total variance with an internal consistency coefficient, $\alpha$ = `r apa(F4.alpha,2,F)`. These five items stemmed from the behavior related to an individual's cellphone usage during the sleep-wakeup time. Lastly, the fifth factor contained six items and explained `r apa(var5,2,F)`% of the total variance. This factor tried to measure an individual's behavior lead by the awareness of light's influence on health. However, this factor showed unsatisfactory internal consistency reliability ($\alpha$ = `r apa(F5.alpha,2,F)`). It is essential to attain a balance between psychometric properties and interpretability of the common themes when exploring the latent structure. As all of the emerged factors are highly interpretable and relevant towards our aim to capture light exposure related behavior, regardless of the apparent low reliability of the two factors, we retain the five-factor solution with 23 items for our confirmatory factor analysis (CFA). Two items showed negative factor-loading (items 44 and 21). Upon inspection, it was understood that these items are negatively correlated to the common theme, and thus in the CFA analysis, we reversed the response code for these two items.

@@

We conducted a categorical confirmatory factor analysis with robust weighted least square (WLSMV) estimator as our response data was in ordinary nature[@desjardinsHandbookEducationalMeasurement2018]. Several indices are suggested to measure model fit. These indices can be categorized as absolute, comparative and parsimony fit indices [@brownConfirmatoryFactorAnalysis2015]. Absolute fit assess the model fit at an absolute level using indices including chi-square test statistics and the standardized root mean square (SRMR).parsimony fit indices including the root mean square error of approximation (RMSEA) considers the number of free parameters in the model to assess the parsimony of the model. Comparative fit indices evaluate the fit of the specified model solution in relation to a more restricted baseline model restricting all covariances among the idicators as zero. Comparative fit index (CFI) and the Tucker Lewis index (TLI) are such two comparative fit indices. Commonly used Model fit guidelines [@huCutoffCriteriaFit1999; @schumacker2004beginner] includes (i) Reporting of chi-square test statistics (A non-significant test statistics is required to reflect model fit) (i) CFI and TLI (CFI/TLI close to .95 or above/ranging between 90-95 and above) (ii) RMSEA (close to .06 or below), (iii) SRMR (close to .08 or below) to estimate the model fit. Table \@ref(tab:tabCfa) summarizes the fit indices of our fitted model. Our fitted model failed to attain an absolute fit estimated by the chi-square test. However, the chi-square test is sensitive to sample size and not recommended to be used as the sole index of absolute model fit [@brownConfirmatoryFactorAnalysis2015]. Another absolute fit index we obtained in our analysis was SRMR which does not work well with categorical data [@RN1272]. Subsequently, we judged the model fit based on the comparative fit indices: CFI, TLI and parsimony fit index-RMSEA. Our fitted model attained acceptable fit (CFI =`r apa(fit.1[5],2,F)`; TLI = `r apa(fit.1[6],2,F)`); RMSEA = `r apa(fit.1[7],2,F)`,\[`r apa(fit.1[8],2,F)`-`r apa(fit.1[9],2,F)`, 90% CI\]) with two imposed equity constrain on item pairs 32-33 and 19-17 . However SRMR value was higher than the guideline (SRMR = `r apa(fit.1[10],2,F)`). Further by allowing one pair of items (30-41) to covary their error variance and discarding two item (item 37 & 26) for very low r-square value, our model attained best fit (CFI =`r apa(fit.2[5],2,F)`; TLI = `r apa(fit.2[6],2,F)`); RMSEA = `r apa(fit.2[7],2,F)`\[`r apa(fit.2[8],2,F)`-`r apa(fit.2[9],2,F)`, 90% CI\]) and SRMR value (SRMR = `r apa(fit.2[10],2,F)`) was also close to the suggestions of @huCutoffCriteriaFit1999. Since reliability coefficient Cronbach's alpha tends to mis-measure reliability for multidimensional construct [@sijtsma2009use] we report Mcdonald's omega(total) as the internal consistency reliability coefficient. Mcdonald's omega(total) is reported to be a satisfactory lower bound reliability coefficient that works for both unidimensional and multidimensional construct [@zinbarg2005cronbach]. Mcdonald's omega(total) for the five factors were `r apa(rel.k[1,1],2,F)`, `r apa(rel.k[5,2],2,F)`, `r apa(rel.k[5,3],2,F)`, `r apa(rel.k[5,4],2,F)`, `r apa(rel.k[5,5],2,F)` respectively. Mcdonald's omega(total) coefficient for the total scale was `r apa(rel.k[5,6],2,F)`.

@@

```{=tex}
\endgroup
```

>>>>

```{r pacman, include=FALSE}
#You need to run this chunck 1X in your system 
#install.packages("pacman")
pacman::p_load(MOTE, tidyverse, psych, lavaan,kableExtra,gt,gtsummary, mirt,likert,kutils,semPlot,semTable,semTools,ggcorrplot,dlookr,qgraph,paran,EFA.MRFA,VIM,DiagrammeR,DiagrammeRsvg,ggplot2,cowplot,rsvg,questionr,magick, simsem,readxl)
pacman::p_install_gh("jthomasmock/gtExtras")
pacman::p_install_gh("crsh/papaja","crsh/citr", "masiraji/tabledown")
pacman::p_install_gh("masiraji/tabledown")

@@

library(papaja) # devtools::install_github("crsh/papaja")
library(lavaan)
library(semPlot) #devtools::install_github('SachaEpskamp/semPlot',  dependencies = T)
library(semTools)
library(MOTE)
library(car)#revercoding 
library(psych)
library(dlookr)
library(plyr)
library(tidyverse)
library(qgraph)
library(kableExtra)
library(paran) # Parallel analysis
library(EFA.MRFA) # Hull method
library(DiagrammeR) #devtools::install_github('rich-iannone/DiagrammeR')
library(DiagrammeRsvg) #For DiagrammeR
library(rsvg) #For DiagrammeR
library(ggcorrplot)
library(semTable)
library(magick)
library(mirt)
library(gtsummary)
library(gt)
library(gtExtras)
library(likert)
library(VIM) #Missing data
library(kutils)
library(simsem)
library(tabledown)
library(readxl)
r_refs("references.bib")
```

@@

```{r}
relevant.scale.data <- readxl::read_excel("Table_raw/relevant_scales.xlsx")
```

```{r}
kable(relevant.scale.data,booktabs = T)
```

@@

-   Sleep disturbance and sleep-related impairment (adult and pediatric versions) [@Bevans.2019; @Buysse.2010; @Forrest.2018; @Harb.2015; @Yu.2011]
-   Sleep duration, timing, and latency, chronotype, social jetlag, time in bed, work/sleep schedule and outdoor light exposure duration (version for adults and adolescents) [@roenneberg2003life]
-   Sleep environment [@Olivier.2016]
-   Meal timing & caffeine consumption [custom items]
-   Light sensitivity (photophobia vs. photophilia) [@Wu.2017]
-   Self-reported pubertal stage (only if younger than 18 years old) [@Petersen.1988]

@@

Age among all participants ranged from `r min(desctable.data$slypos_demographics_age)` years to `r max(desctable.data$slypos_demographics_age)` years [EFA: *min* = `r min(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "1. EFA Sample"])`, *max* = `r max(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "1. EFA Sample"])`; CFA: *min* = `r min(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "2. CFA Sample"])`, *max* = `r max(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "2. CFA Sample"])`], with an overall mean of \~ `r round(mean(desctable.data$slypos_demographics_age), 0)` years of age [Overall: *M* = `r mean(desctable.data$slypos_demographics_age)`, *SD* = `r sd(desctable.data$slypos_demographics_age)`; EFA: *M* = `r mean(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "1. EFA Sample"])`, *SD* = `r sd(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "1. EFA Sample"])`; CFA: *M* = `r mean(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "2. CFA Sample"])`, *SD* = `r sd(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "2. CFA Sample"])`]. In total `r desc_tibble[3,2]` of the participants indicated female sex [EFA: `r desc_tibble[3,3]`; CFA: `r desc_tibble[3,4]`], `r desc_tibble[4,2]` indicated male [EFA: `r desc_tibble[4,3]`; CFA: `r desc_tibble[4,4]`] and `r desc_tibble[5,2]` indicated other sex [EFA: `r desc_tibble[5,3]`, CFA: `r desc_tibble[5,4]`]. Overall, `r desc_tibble[6,2]` [EFA: `r desc_tibble[6,3]`; CFA: `r desc_tibble[6,4]`] participants indicated a gender-variant identity. In a "Yes/No" question regarding native language, `r desc_tibble[7,2]` of respondents [EFA: `r desc_tibble[7,3]`; CFA: `r desc_tibble[7,4]`] indicated to be native English speakers. For their "Occupational Status", more than half of the overall sample reported that they currently work [Overall: `r desc_tibble[9,2]`; EFA: `r desc_tibble[9,3]`; CFA: `r desc_tibble[9,4]`], whereas `r desc_tibble[10,2]` [EFA: `r desc_tibble[10,3]`; CFA: `r desc_tibble[10,4]`] reported that they go to school and `r desc_tibble[11,2]` [EFA: `r desc_tibble[11,3]`; CFA: `r desc_tibble[11,4]`] responded that they do "Neither". With respect to the COVID-19 pandemic we asked participants to indicate their occupational setting during the last four weeks: In the overall sample `r desc_tibble[13,2]` [EFA: `r desc_tibble[13,3]`; CFA: `r desc_tibble[13,4]`] of the participants indicated that they were in a home office/ home schooling setting., while `r desc_tibble[14,2]` overall [EFA: `r desc_tibble[14,3]` ; CFA: `r desc_tibble[14,4]`] reported face-to-face work/schooling. Lastly, `r desc_tibble[15,2]` overall [EFA: `r desc_tibble[15,3]` ; CFA: `r desc_tibble[15,4]`] reported a combination of home- and face-to-face work/schooling, whereas `r desc_tibble[16,2]` overall [EFA: `r desc_tibble[16,3]`; CFA: `r desc_tibble[16,4]`] filled in the "Neither (no work or school, or indication)" response option. We tested all demographic variables in Table 1 for significant group differences between the EFA and CFA sample, applying Wilcoxon rank sum test for the continuous variable "Age" and Pearson's Chi-squared test for all other categorical variables via the gtsummary R package's "add_p" function [@sjoberg2021] . The p-values were corrected for multiple testing applying false discovery rate (FDR) via the "add_q" function of the same package. After p-value (FDR) correction for multiple testing, none of the demographic variables were significantly different between the EFA sample and the CFA sample (all q-values *q* ≥ `r min(desc_tibble$"**q-value**", na.rm=T)` , indicating equivalence).

@@

We used R (version 4.1.0), including several R packages, for our analyses.Initially, our tool have six poin Likert type response scale(0:Does not apply/I don't know; 1:Never, 2:Rarely; 3:Sometimes; 4:Often; 5: Alsways). As our purpose was to capture light exposure related behavior, "Does not apply/I don't know" and "Never" were providing similar information. As such we decided to collapse "Does not apply/I don't know" and "Never" options into one making it a 5 point Likert type response scale. Necessary assumptions of EFA, including sample adequacy, normality assumptions, quality of correlation matrix, were assessed. Our data violated both the univariate and multivariate normality assumptions. Due to these violations and the ordinal nature of our response data, we used a polychoric correlation matrix [@desjardinsHandbookEducationalMeasurement2018] for the EFA. We employed principal axis (PA) as a factor extraction method with varimax rotation. PA is robust to the normality assumption violations [@watkinsStepbyStepGuideExploratory2020]. The obtained latent structure was confirmed by another factor extraction method: the minimum residuals extraction method as well. We used a combination factor identification method including scree plot[@cattellScreeTestNumber1966], Horn's parallel analysis [@hornRationaleTestNumber1965], minimum average partials method[@velicerDeterminingNumberComponents1976], and hull method [@lorenzo-sevaHullMethodSelecting2011] to identify factor numbers. Additionally, to determine the simple structure, we followed the following guidelines recommended by psychometricians (i) no factors with fewer than three items (ii) no factors with a factor loading <0.3 (iii) no items with cross-loading greater than .3 across factors [@bandalosFactorAnalysisExploratory2018] We also conducted psychometric analysis on non-merged response options data (supplementary analysis) and rejected the latent structure obtained as the factors were less interpretable.

@@

Sampling adequacy was checked using Kaiser-Meyer-Olkin (KMO) measures of sampling adequacy[@kaiserIndexFactorialSimplicity1974] . The overall KMO vale for 48 items was `r printnum(KMO$MSA)` which was above the cutoff value (.50) indicating a mediocre sample [@hutchesonMultivariateSocialScientist1999]. Table\@ref(tab:tabDes) summarizes the univariate descriptive statistics for the 48 items. some of the items were skewed with high Kurtosis values. Our data violated both univariate normality (Shapiro-Wilk statistics; [@shapiroAnalysisVarianceTest1965]) and multivariate normality assumptions (Marida's test;[@mardiaMeasuresMultivariateSkewness1970]). Multivariate skew was = `r printnum(mardia$b1p)` (p <0.001) and multivariate kurtosis was = `r printnum(mardia$b2p)` (p <0.001). Due to these violations and ordinal nature of the response data polychoric correlations over Pearson's correlations was chosen [@desjardinsHandbookEducationalMeasurement2018]. Bartlett's test of sphericity [@bartlettNoteMultiplyingFactors1954], $chi^2$ (`r bartlet$df`) = `r apa(bartlet$chisq, 2, T)`, p < .001] indicated the correlations between items are adequate for the EFA. However only `r apa(cor.per, 2,T)`% of the inter-item correlation coefficients were greater than .30. The inter item correlation ranged between `r apa(min.cor,2,F)` to `r apa(max.cor,2,F)`. And the corrected item-total correlations ranged between `r apa(low.r.corec,2,F)` to `r apa(high.r.corec,2,F)`.

@@

With initial 48 items we conducted three rounds of EFA gradually discarded problematic items. (cross-loading items and poor factor loading (<.30) items). Finally, a five-factor EFA solution with 25 items was accepted with low RMSR = 0.08 [@brownConfirmatoryFactorAnalysis2015], all factor-loading higher than .30 and no cross-loading greater than .30. We confirmed this five-factor latent structure using varimax rotation with a minimum residual extraction method (Table\@ref(tab:MinResTab)). Table\@ref(tab:EFATable) displays the factor-loading (structural coefficients) and communality of the items. The absolute value of the factor-loading ranged from `r apa(min.loadings,2,F)` to `r apa(max.loadings, 2, F)` indicating strong coefficients. The commonalities ranged between `r apa(min.com,2,F)` to `r apa(max.com, 2, F)`. However, the histogram of the absolute values of non-redundant residual-correlations Fig\@ref(fig:EFAResiduals) showed `r apa(sumR,0, T)`% correlations greater than the absolute value of .05, indicating a possible under-factoring. [@desjardinsHandbookEducationalMeasurement2018a]. Subsequently, we fitted a six-factor solution. However, a factor emerged with only one salient variable loading in the six-factor solution, thus disqualifying the six-factor solution (Table\@ref(tab:sixFacTab)).

In the five-factor solution, the first factor contained three items and explained `r apa(var1,2,F)`% of the total variance with a satisfactory internal reliability coefficient ($alpha$ = `r apa(F1.alpha,2,F)`). All the items in this factor stemmed from the individual's preference to use blue light filters in different light environments. The second factor contained six items and explained `r apa(var2,2,F)`% of the total variance with a satisfactory internal reliability coefficient ($alpha$ = `r apa(F2.alpha,2,F)`). Items under this factor commonly investigate an individual's hours spent outdoor. The third factor contained five items and explained `r apa(var3,2,F)`% of the total variance. Items under this factor dealt with the specific behaviors pertaining to sleep. The internal consistency reliability coefficient was, $alpha$ = `r apa(F3.alpha,2,F)`. The fourth factor contained five items and explained `r apa(var4,2,F)`% of the total variance with an internal consistency coefficient, $alpha$ = `r apa(F4.alpha,2,F)`. These five items stemmed from the behavior related to an individual's cellphone usage during the sleep-wakeup time. Lastly, the fifth factor contained six items and explained `r apa(var5,2,F)`% of the total variance. This factor tried to measure an individual's behavior lead by the awareness of light's influence on health. However, this factor showed unsatisfactory internal consistency reliability ($alpha$ = `r apa(F5.alpha,2,F)`). It is essential to attain a balance between psychometric properties and interpretability of the common themes when exploring the latent structure. As all of the emerged factors are highly interpretable and relevant towards our aim to capture light exposure related behavior, regardless of the apparent low reliability of the two factors, we retain the five-factor solution with 23 items for our confirmatory factor analysis (CFA). Two items showed negative factor-loading (items 44 and 21). Upon inspection, it was understood that these items are negatively correlated to the common theme, and thus in the CFA analysis, we reversed the response code for these two items.

@@

We conducted a categorical confirmatory factor analysis with robust weighted least square (WLSMV) estimator as our response data was in ordinary nature[@desjardinsHandbookEducationalMeasurement2018]. Several indices are suggested to measure model fit. These indices can be categorized as absolute, comparative and parsimony fit indices [@brownConfirmatoryFactorAnalysis2015]. Absolute fit assess the model fit at an absolute level using indices including chi-square test statistics and the standardized root mean square (SRMR).parsimony fit indices including the root mean square error of approximation (RMSEA) considers the number of free parameters in the model to assess the parsimony of the model. Comparative fit indices evaluate the fit of the specified model solution in relation to a more restricted baseline model restricting all covariances among the idicators as zero. Comparative fit index (CFI) and the Tucker Lewis index (TLI) are such two comparative fit indices. Commonly used Model fit guidelines [@huCutoffCriteriaFit1999; @schumacker2004beginner] includes (i) Reporting of chi-square test statistics (A non-significant test statistics is required to reflect model fit) (i) CFI and TLI (CFI/TLI close to .95 or above/ranging between 90-95 and above) (ii) RMSEA (close to .06 or below), (iii) SRMR (close to .08 or below) to estimate the model fit. Table \@ref(tab:tabCfa) summarizes the fit indices of our fitted model. Our fitted model failed to attain an absolute fit estimated by the chi-square test. However, the chi-square test is sensitive to sample size and not recommended to be used as the sole index of absolute model fit [@brownConfirmatoryFactorAnalysis2015]. Another absolute fit index we obtained in our analysis was SRMR which does not work well with categorical data [@RN1272]. Subsequently, we judged the model fit based on the comparative fit indices: CFI, TLI and parsimony fit index-RMSEA. Our fitted model attained acceptable fit (CFI =`r apa(fit.1[5],2,F)`; TLI = `r apa(fit.1[6],2,F)`); RMSEA = `r apa(fit.1[7],2,F)`,[`r apa(fit.1[8],2,F)`-`r apa(fit.1[9],2,F)`, 90% CI]) with two imposed equity constrain on item pairs 32-33 and 19-17 . However SRMR value was higher than the guideline (SRMR = `r apa(fit.1[10],2,F)`). Further by allowing one pair of items (30-41) to covary their error variance and discarding two item (item 37 & 26) for very low r-square value, our model attained best fit (CFI =`r apa(fit.2[5],2,F)`; TLI = `r apa(fit.2[6],2,F)`); RMSEA = `r apa(fit.2[7],2,F)`[`r apa(fit.2[8],2,F)`-`r apa(fit.2[9],2,F)`, 90% CI]) and SRMR value (SRMR = `r apa(fit.2[10],2,F)`) was also close to the suggestions of @huCutoffCriteriaFit1999. Since reliability coefficient Cronbach's alpha tends to mis-measure reliability for multidimensional construct [@sijtsma2009use] we report Mcdonald's omega(total) as the internal consistency reliability coefficient. Mcdonald's omega(total) is reported to be a satisfactory lower bound reliability coefficient that works for both unidimensional and multidimensional construct [@zinbarg2005cronbach]. Mcdonald's omega(total) for the five factors were `r apa(rel.k[1,1],2,F)`, `r apa(rel.k[5,2],2,F)`, `r apa(rel.k[5,3],2,F)`, `r apa(rel.k[5,4],2,F)`, `r apa(rel.k[5,5],2,F)` respectively. Mcdonald's omega(total) coefficient for the total scale was `r apa(rel.k[5,6],2,F)`.

@@

```{=tex}
\endgroup
```
```{r echo = FALSE, results = 'asis', cache = FALSE}
papaja::render_appendix('Supplymentary.Rmd')
```
